Compositional Certified Resource Bounds\ifshort{\vspace{-.15cm}}{}

### 2014-11-09 paper/main.tex
add paragraph about interaction w/ logic

\begin{figure}
\begin{lstlisting}[numbers=left]
${\color{blue}%
  \mbox{logical state invariant }
  \{ \code{lg} \ge \log_2(h-l) \cdot M_\code{bsearch} \} }$
bsearch(x,l,h_,lg_) {
  if (h-l > 1 && lg > 0) {
    m = h + (h-l)/2;
    _lg--;_ if (a[m]>x) h=m; else l=m;
    tick($M_\code{bsearch}$);
    l = bsearch(x,l,h_,lg_);
    tick($-M_\code{bsearch}$);
  } else return l;
}
\end{lstlisting}
\caption{Example of assisted bound derivation using logical state.
  We use the tick metric to model the stack consumption of the program.
  Before a function call to \code{bsearch}, a tick statement pays for
  the space used by stack frame of the function.  The space is recoverd
  after the call.
  The bound derived automatically is $|[0,\code{lg}]|$.
  }
\label{fig:xmplbsaux}
\end{figure}

@@

Another simple example is given on \pref{fig:xmplbsaux} where
a logarithmic bound on the stack consumption of a binary search
program is proved using logical variable annotations.  Once again,
annotations are underlined in the program text.  In this example,
to ease the proof of equivalence between the annotated program
and the original one, we use the inequality $\code{lg} \ge
\log_2(h-l)\cdot M_\code{bsearch}$ as invariant.  This trick
allows a simpler proof because, when working with integer
arithmetic, it is not always the case that $\log_2(x/2) =
\log_2(x)-1$.

>>>>

\begin{figure}
\begin{lstlisting}[numbers=left]
${\color{blue}%
  \mbox{logical state invariant }
  \{ \code{lg} > \log_2(\code h-\code l) \} }$
bsearch(x,l,h_,lg_) {
  if (h-l > 1 _&& lg > 0_) {
    m = l + (h-l)/2;
    _lg--;_ if (a[m]>x) h=m; else l=m;
    tick($M_\code{bsearch}$);
    l = bsearch(x,l,h_,lg_);
    tick($-M_\code{bsearch}$);
  } else return l;
}
\end{lstlisting}
\caption{Example of assisted bound derivation using logical state.
  We use the tick metric to model the stack consumption of the program.
  Before a function call to \code{bsearch}, a tick statement pays for
  the space used by stack frame of the function.  The space is recoverd
  after the call.  We write $\log_2(x)$ for the integer part of
  logarithm of $x$ in base 2.
  The bound derived automatically is $|[0,\code{lg}]|$.
  }
\label{fig:xmplbsaux}
\end{figure}

@@

Another simple example is given on \pref{fig:xmplbsaux} where
a logarithmic bound on the stack consumption of a binary search
program is proved using logical variable annotations.  Once again,
annotations are underlined in the program text.  In this example,
to ease the proof of equivalence between the annotated program
and the original one, we use the inequality $\code{lg} >
\log_2(\code h-\code l)$ as invariant.  This trick allows a simpler
proof because, when working with integer arithmetic, it is not always
the case that $\log_2(x-x/2) = \log_2(x)-1$ (take 3).

On the other hand, it might be the case that even with annotations, a
resource bound cannot be found automatically.  With existing tools, this
situation requires a full rewrite of the program.  However, in our
framework, it is possible to write a manual proof and combine it with
bounds derived by the automatic analysis.  This interaction is made
formal using our Quantitative Hoare Logic.  This resource aware program
logic is an extension of the one presented in~\cite{veristack14} and
was originally designed to prove the soundness of our automatic
analysis.  To make it suitable for this end, we made it independent
of the resource used (e.g. stack usage, heap usage, clock ticks) and
implemented a smoother handling of the auxiliary state through a shallow
embedding in Coq.  Despite being first a theoritical tool, the logic can
be used directly as in~\cite{veristack14}.  Its judgements are inspired
by typical Hoare logics and have the shape
$$
\Delta \vdash \htriple P S Q
$$
where $\Delta$ is a set of hypothesis on program functions, and $P, Q$
are quantitative assertions, i.e. functions from the program states
to $\Qplusz \cup \{\infty\}$.  The semantics of triples and the rules
of the logic are described in further details in the extended version of
this paper~\cite{anon_extended}.  Because our automated tool produces
proofs of the resource bounds in the Quantitative Hoare Logic, it can
be used as an assistant when deriving a complex bound by hand in the
logic.  The interaction is guaranteed to be sound by the soundness
theorem of our Quantitative Hoare Logic.
